Speculation in foreign exchange: noise or information

### 2015-07-16 Knitr/NewPaper.Rnw
This paper has been completed NewPaper.Rnw.  However, there is some issue with the appendix. I think that is to do with the table.

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{appendix}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
An event study is used to assess the views of Keynes and Friedman on speculation. Speculative extremes are ranked by intensity of sentiment and weight of speculative activity. A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment; the weight of positions is based on US regulatory data.  The extremes say little about the future in either case, suggesting that speculation is more than just random noise and supporting the view there is some informational content that is being passed through to the price with speculative activity. 
\end{abstract}
\section*{Introduction}
There are a range of ideas about the way that speculation can affect financial markets.  Towards one end of this spectrum is the view of Keynes from Chapter 12 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p. 101]{Keynes1936}. Here speculation is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that does not emerge randomly, but feeds off itself in an evolution of collective sentiment that encourages trends in asset prices.

@@

One key element of the microstructure literature involves distinguishing between uninformed or noise-traders and those who have information. The \citet{Kyle1985Continuous} model has market-makers receiving orders from the informed and the uninformed.  On the assumption that the uniformed arrive in a random fashion without any reason to trade, the weight of orders implies something about the activity of the informed and causes prices to change in the process of price discovery.  The bid-ask spread is the payment that the market-maker receives in return for taking the risk of meeting an informed trader.  Ultimately, the uninformed traders make payments to the informed traders with the mark-makers taking a cut for the risk and the provision of exchange.  See also \citet{glosten1985bid}.

In the \emph{microstructure} literature, the term noise-trader is often interchanged with \emph{liquidity-trader}.  They do not usually distinguish between institutional and psychological reasons for the lack of attention to fundamental information. Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value. This may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy.

\citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. In this case, uninformed speculation provides a rational and an opportunity for the trading of financial securities. If all economic agents know the fundamental value of a security how can some be buying and some be selling? There must be different opinions.  Why will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information? With speculative noise-traders, the informed can take advantage of their information.  This creates a justification for acquiring information. Speculative noise-traders can therefore create a gap between the price and the fundamental value that will encourage informed traders to pay the cost of acquiring information. The divergence between this price and fundamental value is thereby essential to encourage the process of price discovery. 
%
 \citet[p.529]{BlackNoise} suggested that this combination was essential to the smooth function of financial markets. It is ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value. This begs the question: are speculators trading on noise or are they informed?

  Speculation is also an important part of the argument for why prices should tend towards a level that consistent with \emph{the rational expectations equilibrium} (RER). This is the level that is consistent with rational expectations.  In the original version of rational expectations,  \citet{muth1961Rational} uses commodity speculation as the means by which prices are driven towards the level where expectations are model-consistent. These are of Friedman-type speculators as they base their expectations on their knowledge of the underlying model and they are the mechanisem through which equilibrium is achieved.

However, though equating noise and noise-trading as a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, it may also be an over-simplification that is inappropriate when looking for more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speuclative activity.

The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.  See \citet{BrunnermeierLiquidity} and \citet{Gorton} for an overview of the institutional literature.  Psychological explanations revolve around the common panic, the group-thinking and the tendency to read private information from the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than actual underlying, fundamental value \citet[p.101]{Keynes1936}.

\subsection{Empirical work}
Empirical assessment of speculation has been limited by the scarcity of data.  It is not very easy to distinguish between speculators and other types of investors.  It is even harder to distinguish those that may be noise-traders from those assumed to be informed.  One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.  For example speculation can be identified with momentum trading. It is also possible to assume that certain investors, such as investment banks and hedge funds, are informed while other trade without information. The latter group may include day-traders or private individuals.  
It has also been suggested that speculative techniques can change over time.  For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.  In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation.

@@

\subsection{Why this is important}
The Keynesian view of speculation is that it is an addition to market noise that makes it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value. As such, it exacerbates the tendency towards putting too much emphasis on the immediate and works against longer-term and strategic thinking.  This description provides an explanation for Shiller's observation that stock prices are more volatile than the underlying earnings information on which valuation should be based \citet{ShillerVol}. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.

The nature of speculation and the role of noise-traders has become increasingly important since the global financial crisis (GFC).  There are a number of contemporary questions where the role and nature of speculation is at the centre of the debate.  For example, regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial as they reduce risk and encourage more long-term attitudes towards investment.  However, where these measures reduce liquidity, higher risk and an increase in other costs may provide an countervailing force.

@@

\section{Methodology}
The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the foreign exchange market. This will be used to gain a fuller understanding of speculative behaviour.   The analysis will involve three steps:  assessement of the price implication of the noise-trader model;  measurement of speculative activity in the foreign exchange market; finally, assessement of whether extreme speculative market activity is consistent with the noise-trader model.  If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.

The \citet{Delong1990noise} noise-trader model used in this study. This is a derivative of the \citet{Samuelson1958} overlapping generations model with two-period agents making a portfolio choice between two assets: a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; a risky asset $(u)$ also pays the same dividend but is not in elastic supply.  There are two types of investor: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.  In the absence of noise-traders, the assets are substitutes.  Each trader chooses their portfolio in the first period $(t)$ to maximise their utility on the basis of their belief about the price of the unsafe asset in the second period $(t + 1)$.  The representative noise-trader mistakes the price of the risky asset by a random variable

@@

Maximising utility with respect to holdings of the risky asset and iterating backward to solve for the current value,  the price becomes the discounted value of the future dividends, the expected capital gain and the general misperception of noise-traders less risk.

@@

The following four effects are noted by \citet[pp. 14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} as the speculative noise-traders drive prices towards their misperceived value; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the noise-trader risk caused by the difference between their perception and the actual value, it is an added risk that is caused by the variability of noise-trader belief as noise-traders could drive the price even further from its fundamental value. In this model, the more numerous the noise-traders, the more volatile prices.

The Friedman-effect and the Price-pressure are the effects that are studied here.  Their combination suggests that when the weight of noise-traders is high or the times when they have particularly strong opinions will tend to be those when there is the greatest risk of reversal.  The hold-more effect suggests that the weight of speculative involvement in the market will tend to increase as a trend.  Create-space effect would suggest that when the weight of noise-traders is high, there is increased risk.  %this can be assessed.

@@

A 1000 period prices simulation is created for each model and the returns for these series are calculated as the log change in price.  Those periods when misperception-perception or the weight of speculators is \emph{extreme} are identified as those that are at or above the ninety-fifth percentile.  These extremes are the \emph{event} for an event study. The R package \emph{eventstudies} created by \citet{eventstudies} is used for the data manipulation.  Descriptive statistics for the whole series of returns and for those when either the extent of their misperception or the weight of noise-traders is extreme are presented in Table \ref{tabref:MC1}.  The process is then repeated for one thousand simulations of each model so that confidence intervals can be established for the estimates of the descriptive statistics.

@@

\subsection{Measuring the intensity of speculative sentiment}
The measurement of speculative intensity must identify how strongly speculators believe that the market will move in a particular direction. In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.%\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.  However, these are disparate and, in many cases, unsystematic, and it is not clear whether they measure speculative sentiment or investor opinion more generally.   The Wall Street Journal runs an annual survey of investor opinion, but this is focused mostly on US equities.  The \emph{Economist} magazine also solicits opinion, but in a relatively unstructured and \emph{ad hoc} fashion.  In the foreign exchange market, there is a broad set of market forecasts that was collected by the financial services company \emph{MMS International} (formerly Money Market Services) which polled a broad range of financial market participants on key exchange rates for one and four week forecast horizons. Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.  Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citet gc{FrootFrankelFDB} and \citep{FrankelFroot1987}.}
Risk-reversal skew is the relative price of currency put and call options for the same strike price.  The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same.  While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.   The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.   The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.   This model allows the market practice of pricing options from the volatility implied by the model.  In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.

This study uses a unique dataset of \emph{risk-reversals} for eleven major currency pairs.  This is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  A call option is the right to buy and a put option is the right to sell.  The delta of an option measures the effect of a change in the spot price on the value of the option.  This can range from zero for deeply out-of-the-money options to unity for deeply in the money strikes.  The at-the-money delta is 50 as there is a 50:50 chance of the option being exercised.  The benchmark for option risk-reversals is 25 delta.  This is a measure of how far the option is out of the money or how far the strike is from the forward rate.

If the exchange rate trajectory is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.  Any deviation from this equivalence is an indication that a deviation from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.  The dealer pays 0.15\%.  Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\% \citet{Global}. }  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity can be taken as an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls or, in this case, as an indication of market sentiment.  The larger the absolute level of the risk-reversal skew, the greater the relative price and, in equilibrium, the greater the belief, expressed through the action of market participants or the inventory of market-makers, that there will be a deviation from the forward rate.   If extreme deviations or speculations are noise they should presage a sharp reversal.  If speculation is informed, this is part of the price discovery process and extremes may be followed by a random walk.

@@

Table \ref{tabref:DS2} gives the descriptive statistics for the risk-reversal data.   For a neutral risk-reversal, a reading of zero, there would be no price preference for calls or puts at the similar delta.  The figures show that, in contrast to the average returns, there are often average skews that are different from zero.  The most significant of these are the positive bias for EUR-GBP and a negative bias for USD-JPY.  The latter may be related to hedging activity by Japanese exporters.  While this could suggest that implied volatility is an imperfect measure of speculative activity, note that extreme readings are based on quantiles of the distribution and therefore a discrete adjustment of this distribution to reflect the hedging power should not change the timing of signals measured by extremes.

@@

\subsection{Analysis of events and abnormal returns}
The aim of the study is to assess the effect of extreme speculative positions on future returns to ascertain the informational content of these extremes.  Therefore, the events that will be used to assess abnormal returns are the extreme readings on sentiment and the extreme presence of speculators.   Any reading above the ninety-ninth percentile or below the first percentile is deemed to be extreme.  For the risk reversal skew, as a secondary test, a reading above the ninety-fifth and below the fifth percentile is also established as an event. The \citet{eventstudies} R package was used to manage the data.  The cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) were recorded.  These cumulations are just the summation of the continuously compounded returns over the appropriate period.

The date of the events, 1st, 5th, 95th or 99th percentile, reading of the risk-reversal skew or the level of speculative positions,  are identified and the exchange rate returns for the window around the event are aligned so that the average for each event and each event day can be calculated.  For example, in Table \ref{tabref:RR1} the analysis of the event that the risk-reversal skew for the EUR vs the US dollar is an extreme low (measured as being equal to the 5th percentile or lower), all 106 cases of this extreme for the period under study are aligned and the average cumulative returns for the whole window (CARW - one day before and after the event in column four) and the post-event window (CARA - the event and the day after the event) are calculated.   A \emph{bootstrap} is conducted.  This means that a sample of size 106 (in this case), with replacement, is drawn from each of the event days and new means are calculated for the CARW and CARA.   If 95\% of the means that are calculated after drawing random samples from the array of windows are above or below zero, it is assumed that the figure is statistically different from zero, and this is represented by an asterisk in the table.

@@

What is more interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.  The Friedman-effect would suggest a reversal after an extreme.   This information is provided in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}. A reversal is indicated by a significant negative reading for highs or positive reading for lows.   Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.  Of the eighteen cases using the 95\% th percentile to identify an extreme, ten show a continuation, rather than a reversal, that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).  For the more extreme event when the 99\% the percentile is used to measure the extreme, there is an even more mixed picture.  Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant.

>>>>

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{appendix}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
An event study is used to assess the views of Keynes and Friedman on speculation. Speculative extremes are ranked by intensity of sentiment and weight of speculative activity. A unique dataset of risk-reversal skew on option prices is used to measure the intensity of speculative sentiment; the weight of positions is based on US regulatory data.  The extremes say little about the future in either case, suggesting that speculation is more than just random noise and supporting the view there is some informational content that is being passed through to the price with speculative activity. 
\end{abstract}
\section*{Introduction}
There are a range of ideas about the way that speculation can affect financial markets.  Towards one end of this spectrum is the view of  \citet[p. 101]{Keynes1936} in Chapter 12 of \emph{The General Theory} where he identified speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''. Here speculation is a problem that, at its most extreme, will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that does not emerge randomly, but feeds off itself in an evolution of collective sentiment that encourages trends in asset prices.

@@

One key element of the microstructure literature involves distinguishing between uninformed or noise-traders and those who have information. The \citet{Kyle1985Continuous} model has market-makers receiving orders from the informed and the uninformed.  On the assumption that the uniformed arrive in a random fashion without any reason to trade, the weight of orders implies something about the activity of the informed and causes prices to change in the process of price discovery.  The bid-ask spread is the payment that the market-maker receives in return for taking the risk of meeting an informed trader.  Ultimately, the uninformed traders make payments to the informed traders with the market-makers taking a cut for the risk and the provision of exchange.  See also \citet{glosten1985bid}.

In the \emph{microstructure} literature, the term noise-trader is often interchanged with \emph{liquidity-trader}.  There may be institutional and psychological reasons for the lack of attention to fundamental information. Nonetheless, liquidity tends to refer to institutional explanations for trading without reference to fundamental value. This may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy. Noise is more psychological.

\citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. In this case, uninformed speculation provides a rational and an opportunity for the trading of financial securities. There must be different opinions.  Why will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information? With speculative noise-traders, prepared to take the other side of the transaction, the informed can take advantage of their information.  This creates a justification for acquiring information. Speculative noise-traders can therefore create a gap between the price and the fundamental value that will encourage informed traders to pay the cost of acquiring information.   \citet[p.529]{BlackNoise} suggested that this combination was essential to the smooth function of financial markets. It is ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value. This begs the question: are speculators trading on noise or are they informed?

  Speculation is also an important part of the argument for why prices should tend towards a level that consistent with \emph{the rational expectations equilibrium} (RER). This is the level that is consistent with rational expectations.  In the original version of rational expectations,  \citet{muth1961Rational} uses commodity speculation as the means by which prices are driven towards the level where expectations are model-consistent. These are of Friedman-type speculators as they base their expectations on their knowledge of the underlying model and they are the mechanism through which equilibrium is achieved.

However, though equating noise and noise-trading as a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, it may also be an over-simplification that is inappropriate when looking for more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.

  The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.  See \citet{BrunnermeierLiquidity} and \citet{Gorton} for an overview of the institutional literature.  Psychological explanations revolve around the common panic, the group-thinking and the tendency to read private information into the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than to form an independent view of the actual underlying, fundamental value \citet[p.101]{Keynes1936}.

\subsection{Empirical work}
Empirical assessment of speculation has been limited by the scarcity of data.  It is not very easy to distinguish between speculators and other types of investors.  It is even harder to distinguish those that may be noise-traders from those assumed to be informed.  One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.  For example speculation can be identified with momentum trading. It is also possible to assume that certain investors, such as investment banks and hedge funds, are informed while other trade without information. The latter group may include day-traders or private individuals.  
It has also been suggested that speculative techniques can change over time.  For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.  In this way, speculation can be be associated with sentiment and measures of sentiment are then used as ways of measuring speculation.

@@

\subsection{Why this is important}
The Keynesian view of speculation is that it is an addition to market noise that makes it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value. As such, it exacerbates the tendency towards putting too much emphasis on the immediate and works against longer-term and strategic thinking.  This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.

The nature of speculation and the role of noise-traders has become increasingly important since the global financial crisis (GFC).  There are a number of contemporary questions where the role and nature of speculation is at the centre of the debate.  For example, regulations or taxes that reduce the amount of speculation have generally been considered to be beneficial as they limit risk-taking and encourage more long-term attitudes towards investment.  However, where these measures reduce liquidity, higher risk and an increase in other costs may provide an countervailing force.

@@

\section{Methodology}
The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the foreign exchange market.  The analysis will involve three steps:  assessment of the price implication of the noise-trader model;  measurement of speculative activity in the foreign exchange market; finally, assessment of whether extreme speculative market activity is consistent with the underlying model that is used.  If the speculative activity in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is closer to the Keynesian view than that of Friedman.

The \citet{Delong1990noise} noise-trader model used in this study. This is a derivative of the \citet{Samuelson1958} overlapping generations model with two-period agents making a portfolio choice between two assets: a safe asset $(s)$ which pays a set dividend $(d)$ and is in perfectly elastic supply; a risky asset $(u)$ which also pays the same dividend but is not in elastic supply.  There are two types of investor: sophisticated traders $(i)$ who know the fundamental value of the risky assets and noise-traders $(n)$ who misperceive the fundamental value.  In the absence of noise-traders, the assets are substitutes.  Each trader chooses their portfolio in the first period $(t)$ to maximise their utility on the basis of their belief about the price of the unsafe asset in the second period $(t + 1)$.  The representative noise-trader mistakes the price of the risky asset by a random variable

@@

Maximising utility with respect to holdings of the sage and risky assets for the informed and uninformed traders and iterating backward to solve for the current value,  the price becomes the discounted value of the future dividends, the expected capital gain and the general misperception of noise-traders less risk.

@@

The following four effects are noted by \citet[pp. 14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} is the \emph{noise-trader risk} as the speculative noise-traders drive prices towards their misperceived value; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the difference between noise-trader perception and the actual value, it creates the opportunity for informed traders to profit from their information. In this model, the more numerous the noise-traders, the more volatile prices.

The Friedman-effect and the Price-pressure are the effects that are studied here.  Their combination suggests that when the weight of noise-traders is high or the times when they have particularly strong opinions will tend to be those when there is the greatest risk of reversal.  The hold-more effect suggests that the weight of speculative involvement in the market will tend to increase as a trend.  Create-space effect would suggest that when the weight of noise-traders is high and extreme, there is greater incentive and opportunity for the informed to take advantage of the divergence between fundamental value and price.

@@

A 1000 period prices simulation is created for each model and the returns for these series are calculated as the log change in price.  Those periods when misperception-perception or the weight of speculators is \emph{extreme} are identified as those that are at or above the ninety-fifth percentile.  These extremes are the \emph{event} for an event study. Descriptive statistics for the whole series of returns and for those when either the extent of their misperception or the weight of noise-traders is extreme are presented in Table \ref{tabref:MC1}.  The process is then repeated for one thousand simulations of each model so that confidence intervals can be established for the estimates of the descriptive statistics.

@@

\subsection{Measuring the intensity of speculative sentiment}
The measurement of speculative intensity must identify how strongly speculators believe that the market will move in a particular direction. In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.%\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.  However, these are disparate and, in many cases, unsystematic, and it is not clear whether they measure speculative sentiment or investor opinion more generally.   The Wall Street Journal runs an annual survey of investor opinion, but this is focused mostly on US equities.  The \emph{Economist} magazine also solicits opinion, but in a relatively unstructured and \emph{ad hoc} fashion.  In the foreign exchange market, there is a broad set of market forecasts that was collected by the financial services company \emph{MMS International} (formerly Money Market Services) which polled a broad range of financial market participants on key exchange rates for one and four week forecast horizons. Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.  Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citet gc{FrootFrankelFDB} and \citep{FrankelFroot1987}.}
Risk-reversal skew is the relative price of currency put and call options for the same strike price.  The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not\footnote{While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.   The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.   The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.   This model allows the market practice of pricing options from the volatility implied by the model.  In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters}.

This study uses a unique dataset of \emph{risk-reversals} for eleven major currency pairs.  It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  A call option is the right to buy and a put option is the right to sell.  The delta of an option measures the effect of a change in the spot price on the value of the option.  This can range from zero for deeply out-of-the-money options to unity for deeply in the money strikes.  The at-the-money delta is 50 as there is a 50:50 chance of the option being exercised.  The benchmark for option risk-reversals is 25 delta.  This is a measure of how far the option is out of the money or how far the strike is from the forward rate.

If the exchange rate trajectory is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.  Any deviation from this equivalence is an indication that a deviation from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.  The dealer pays 0.15\%.  Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\% \citet{Global}. }  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment.  The larger the absolute level of the risk-reversal skew, the greater the relative price and, in equilibrium, the more intense the belief, expressed through the action of market participants or in the inventory of market-makers, that there will be a deviation from the forward rate.   If sentiment is based   on noise rather than information, extremes should presage a sharp reversal; if speculative sentiment is informed, this is part of the price discovery process and extremes may be followed by a random walk.

@@

Table \ref{tabref:DS2} gives the descriptive statistics for the risk-reversal data.   For a neutral risk-reversal, a reading of zero, there would be no price preference for calls or puts for the same delta.  The figures show that, in contrast to the average returns, there are often average skews that are different from zero.  The most significant of these are the positive bias for EUR-GBP and a negative bias for USD-JPY.  The latter may be related to hedging activity by Japanese exporters.  While this could suggest that implied volatility is an imperfect measure of speculative activity, note that extreme readings are based on quantiles of the distribution and therefore a discrete adjustment of this distribution to reflect the hedging power should not change the timing of signals measured by extremes.

@@

\subsection{Analysis of events and abnormal returns}
The aim of the study is to assess the effect of extreme speculative positions on future returns to ascertain the informational content of these extremes.  Therefore, the events that will be used to assess abnormal returns are the extreme readings on sentiment and the extreme presence of speculators.   Any reading above the ninety-ninth percentile or below the first percentile is deemed to be extreme.  For the risk reversal skew, as a secondary test, a reading above the ninety-fifth and below the fifth percentile is also established as an event. The  R package \emph{eventstudies} \citet{eventstudies} was used to manage the data.  The cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) were recorded.  These cumulations are just the summation of the continuously compounded returns over the appropriate period.

The date of the events, 1st, 5th, 95th or 99th percentile, reading of the risk-reversal skew or the level of speculative positions,  are identified and the exchange rate returns for the window around the event are aligned so that the average for each event and each event day can be calculated.  For example, in Table \ref{tabref:RR1} the analysis of the event that the risk-reversal skew for the EUR vs the US dollar is an extreme low (measured as being equal to the 5th percentile or lower), all 106 cases of this extreme for the period under study are aligned and the average cumulative returns for the whole window (CARW - one day before and after the event in column four) and the post-event window (CARA - the event and the day after the event) are calculated.   A \emph{bootstrap} is conducted.  This means that a sample of size 106 (in this case), with replacement, is drawn from each of the event days and new means are calculated for the CARW and CARA.   If 95\% of the means that are calculated after drawing random samples from the array of windows are above or below zero, it is asserted that the figure is statistically different from zero, and this is represented by an asterisk in the table.

@@

What is more interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.  If this speculation is noise, Friedman's informed traders should take advantage of that and a reversal should be seen.   This information is provided in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}. A reversal is indicated by a significant negative reading for highs or positive reading for lows.   Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.  Of the eighteen cases using the 95\% th percentile to identify an extreme, ten show a continuation, rather than a reversal, that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).  For the more extreme event when the 99\% the percentile is used to measure the extreme, there is even more ambiguity.  Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant.
